### 2013-10-24 PositionPaper/relwork.tex

\begin{itemize}
\item security concerns - SLA 
\item data integration (cf. travaux Gonzalez) 
\end{itemize}