### 2013-11-24 PositionPaper/conclusions.tex
With conclusions

This paper introduces the challenge of integrating data from distributed data services deployed on different cloud providers guided by service level agreement. The data integratio problem is stated as a continuos data provision problem that has ans associated economic cost and that uses automatic learning techniques for ensuring different qualities of delivered data (fresh, precise, partial).

Current big data settings impose today to consider SLA and different data delivery models. We believe that given the volume and the complexity of query evaluation that includes steps that imply greedy computations, it is important to combine and revisit well-knowns solutions and adapted to this contexts. We are currently developing the strategies and algorithms sketched here applied to energy consumption applications as the one described in the paper and also to elections and political campaign data integration in order to guide decision making on campaign strategies.