### 2013-11-29 PositionPaper/relwork.tex
no message

%\begin{itemize}
%\item Cloud computing represents a novel on-demand computing approach where resources are provided in compliance to a set of predefined non-functional properties specified and negotiated by means of Service Level Agreements (SLAs).  SLA currently exploited in the cloud allows service providers and cloud client to fix the resource level that should be used either by a service for the service provider or by the client in the case of service invokation. SLA could concern all types of the services on the cloud (IAAS, PAAS or SAAS). One of the weakness of SLA for the client is that they express low level clauses like storage amount use or virtualization level. In fact when a client launches a task on the cloud, she has unlikely Q \& S requirement rather than low level features. Recently we notice the existing of a lot of works with the objective to fill the gap.  Emeakaroha and al propose in \cite{5547150}  a cloud component that acts autonomously based on low level  monitored features after analysing them towards high level clauses expressed by the user. Moreover, another difficulty to enforce an SLA is to measure and identify, starting from a high level SLA clause, how could it be declined at different layers in the cloud. Therefore in \cite{Dastjerdi:2012:DOA:2275356.2275360} the authors describe a semanticSLA which can be understood by all parties including providers,
%requestors, and monitoring services. One of the major objectives in the cloud is to anticipate SLA violation and to assess SLA failure cascading on violation detection. \cite{Dastjerdi:2012:DOA:2275356.2275360} proposes an SLA dependency modeling using Web Service Modeling Ontology (WSMO) to build a knowledge database.  \cite{5547150} proposes to anticipate failure by analysing the monitored feature and to act by anticipation. in \cite{5614035}, the authors propose LAYSI, a layered solution that minimizes user interactions with the system and prevents violations of agreed SLAs. 
%On the other hand SLA contracts do not cover all client requirement. There is still a lack in some domain.....(to be completed)

%\item some

The advent of cloud computing imposed a new model for resource consuming that is not concerned
 with resource availability but with the \textit{technical and economic} conditions to be fulfilled in order to access a potentially infinite set of resources. 
Service Level Agreements (SLA)~\cite{SLA} can be used in the context of cloud computing in order to establish those conditions. The deployment of SLAs relies on two principles: (i) The negotiation of conditions, which are statically agreed between the parts and (ii) The monitoring of these conditions during the use of cloud resources in order to detect SLA contract violation.

%The agreed conditions should be explicit and quantifiable.
%They should be part of the contract between the client and the service provider.
%Their quantifiable nature makes it possible for the conditions to appear as non-functional requirement to be imposed to each individual service.
%SLA could concern all types of services on the cloud (IaaS, PaaS or SaaS).

%SLA compliance may be ensured by construction or may be monitored during the execution of the service.
%Monitoring SLAs is normally part of the infrastructure provided by the cloud.
%Service monitoring may incur in extra processing costs.

%SLA currently exploited in the cloud allows service providers and cloud clients to describe the amount of resources to be simultaneously used and also the conditions to use them.
%This limit may apply both for the service provider and the client (in the case of service invocation).

Negotiating a SLA may be a difficult task.
Indeed, SLA are normally expressed in terms of \textit{low level} concepts, like size of the used storage or degree of virtualization. 
On the client side, the interest is normally on \textit{higher level} concepts such as QoS requirements.
Some recent proposals address this problem.  
For instance,~\cite{5547150} propose a bottom-up approach, where a cloud component that acts autonomously based on low level monitored features after analysing them towards high level clauses expressed by the user. 
  \cite{Dastjerdi:2012:DOA:2275356.2275360} describe a \textit{Semantic SLA} which can be understood by all parties including providers, requestors, and monitoring services.
Their goal is, starting from a high level SLA, to measure and identify how the SLA be declined at different layers in the cloud. In the domain of cloud data service, Ortiz and al  \cite{Ortiz:2013:VPS:2486767.2486772}
  proposes a set of templates to the user of cloud data services, each specifying the query type that can be executed with some tread-off in time and corresponding cost.  The client has just to choose among proposed tempate the one that corresponds the best to her requirements. \cite{6141307} proposes a set of matching algorithms to identify compatible cloud provider for a given requirements by matching SLA parameters. In this work, SLA parameters of the cloud are defined in a cloud capable model and application's requirements are defined in an application requirement model. Those models specified firstly using RDF, are converted into graph structures  
From whom an Induced Propagation Graph is calculate to establish a correspondence between the cloud capability model and the application requirement model.

%{\color{red}
%One of the major objectives in the cloud is to anticipate SLA violation and to assess SLA failure cascading on violation detection. 
%\cite{Dastjerdi:2012:DOA:2275356.2275360} proposes an SLA dependency modeling using Web Service Modeling Ontology (WSMO) to build a knowledge database.  
%\cite{5547150} proposes to anticipate failure by analysing the monitored feature and to act by anticipation. 
%In \cite{5614035}, the authors propose LAYSI, a layered solution that minimizes user interactions with the system and prevents violations of agreed SLAs. 
%}

Some recent research works on SLA are devoted to the SLA extension to security concerns. 
\cite{6274042} propose an extension of the WSAgreement, initially developed by the GRAAP
working group. 
The extension allows security constraints to be expressed over the service description
terms (SDTs) and the service level objectives (SLOs) of the SLA. 
Consequently, this leads to a certain kind of security expression interoperability and lets the client to be able to compare security level of different cloud service provider. 
\cite{LunaGarcia:2012:BCS:2381913.2381932} focuses on how to build a SEcSLA Template starting from gathering then categorizing a set of security statements using a semantic tool. 
The designed template is then used both to express user security requirements and cloud service provider security provisions.
Finally, some works tend to deal with SLA violation anticipation and SLA failure cascading on violation detection.  \cite{Dastjerdi:2012:DOA:2275356.2275360}  proposes an SLA dependency modeling using Web Service Modeling Ontology (WSMO) to build a knowledge database. \cite{5614035} proposes to anticipate failure by analysing the monitored feature and to act by anticipation. in
\cite{5614035}, the authors propose LAYSI, a layered solution that minimizes user interactions with the system and prevents violations of agreed SLAs.

%=====================================

%Another issue in the SLA management domain is how to let cloud service provider offer fit the best user requirements. In fact, current SLA are only resource oriented and lacks in expressing client requirement in service quality and characteristics. Plenty of works tends to fill the gap. First by proposing predefined templates;.

%  In \cite{LunaGarcia:2012:BCS:2381913.2381932,}, the authors propose to do a mapping of both user SecSLA requirements and CSP SecSLA provisions on a set of Quantitative Policy Tree(QPT), where atomic capabilities are mapped on leaves and intermediate represents coarse provisions with some logical operators (AND, OR). The mapping allows to give a quantitative benchmarking to each CSP offer with respect to user requirements that lets the user be able to make a choice. For more experienced users, \cite{6313668} proposes an extensible specification grammar that allows to express user and application-specific requirements in using a cloud data service. This allows to customize the SLA management at the cloud side. Emeakaroha and al propose in \cite{5547150}  a cloud component that acts autonomously based on low level monitored features after analysing them towards high level clauses expressed by the user. Moreover, another difficulty to enforce an SLA is to measure and identify, starting from a high level SLA clause, how could it be declined at different layers in the cloud. Therefore in  \cite{Dastjerdi:2012:DOA:2275356.2275360}  the authors describe a semantic- SLA which can be understood by all parties including providers,requestors, and monitoring services. 
% \cite{6141307} proposes a batterie of matching algorithms to identify compatible cloud provider for a given requirements by matching SLA parameters.
%In this work, SLA parameters of the cloud are defined in a cloud capable model and application's requirements are defined
%in an application requirement model. Those models specified firstly using RDF, are converted into graph structures using Jena APIs.
%From those graphs a Pairwise Connectivity Graph is defined and an Induced Propagation Graph is calculated.
%Pairwise Connectivity Graph is constituted by nodes called mappairs established when source and target nodes from initial graphs have the same property edge.
%For the Propagation Graph each edge is given a weight. Finally, a Similarity propagation graph is constructed based on an initial mapping between the RDF models, where mapping nodes are compared (subclass, superclass, equivalence relation) and the mapping is used in a similarity flooding algorithm which is filtered out to only the instance node pairs that have similarity value greater than zero.
%

 % \item data integration (cf. travaux Gonzalez) 
%\end{itemize}